---
title: "STAT 6820: Assignment 6"
author: "Hyrum Hansen"
date: '2023-03-30'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(matlib)
library(xtable)
library(stats)
library(dplyr)
library(tidyverse)
library(ggplot2)
```

```{r}

##########################
## FUNCTION DEFINITIONS ##
##########################

# This function computes the D score for a given matrix
compute_D <- function(X, n_factors, n_trials){
  
  # Convert to matrix if necessary
  if(is.vector(X)){
    X <- matrix(X, n_trials, n_factors)
  }
  
  intercept <- rep(1, nrow(X))
  interaction <- X[,2]*X[,3]
  model_matrix <- cbind(intercept, X, interaction)
  
  # Then compute the D score for the candidate design
  return(det(t(model_matrix)%*%model_matrix))
}

# This function computes the A score for a given matrix
compute_A <- function(X, n_factors, n_trials){
  
  # Convert to matrix if necessary
  if(is.vector(X)){
    X <- matrix(X, n_trials, n_factors)
  }
  
  intercept <- rep(1, nrow(X))
  interaction <- X[,2]*X[,3]
  model_matrix <- cbind(intercept, X, interaction)
  
  # Then compute the A criteria for the candidate design
  return(sum(1/eigen(t(model_matrix)%*%model_matrix)$values))
}

# This function computes the I score for the 2 factor setting
compute_I_2 <- function(X){
  
  # We need to get the moment matrix
  row1 <- c(4, 0, 0, 0, 4/3, 4/3)
  row2 <- c(0, 4/3, 0, 0, 0, 0)
  row3 <- c(0, 0, 4/3, 0, 0, 0)
  row4 <- c(0, 0, 0, 4/9, 0, 0)
  row5 <- c(4/3, 0, 0, 0, 4/5, 4/9)
  row6 <- c(4/3, 0, 0, 0, 4/9, 4/5)
  moment_matrix <- matrix(c(row1, row2, row3, row4, row5, row6), byrow = TRUE, nrow = 6, ncol = 6)

  intercept <- rep(1, nrow(X))
  interaction <- X[,1]*X[,2]
  x1_sq <- X[,1]^2
  x2_sq <- X[,2]^2
  model_matrix <- cbind(intercept, X, interaction, x1_sq, x2_sq)
  return((1/4)*sum(diag(inv(t(model_matrix)%*%model_matrix))))
}



# Following the example of Peter Goos and Brad Jones, this function
# uses a maximization perspective to find the optimal model matrix.
coordinate_exchange <- function(n_factors, n_trials, npasses, testvec, crit){

  # Instantiate a matrix of random values
  n_rows = n_trials
  n_cols = n_factors
  X <- matrix(c(runif(n_rows*n_cols, min=-1, max=1)), nrow=n_rows, ncol=n_cols)
  # Compute the D optimality score until the number of trials has been
  # reached or until there is no more improvement
  trial <- 0
  testvalues <- testvec

  while(trial < npasses){

    candidate_matrices = lapply(1:length(testvalues),
                                matrix, data= NA, nrow=n_rows, ncol=n_cols)
    Xprevious <- X
    # This thingy exchanges the coordinates
    for(i in 1:n_rows){
      for(j in 1:n_cols){
        for(val in 1:length(testvalues)){

          xtest <- X
          xtest[i, j] <- testvalues[val]
          candidate_matrices[val][[1]] <- xtest

        }
        if(crit == "D"){
          # This thingy determines which matrix in the list of candidate matrices
          # has a higher D score
          d_scores <- unlist(lapply(candidate_matrices, compute_D))
          max_index <- which.max(d_scores)
          
          # Now update if the matrix with one element swapped has a greater D score
          if(compute_D(X) < compute_D(candidate_matrices[[max_index]])){
            X <- candidate_matrices[[max_index]]
          }
        }
        
        if(crit == "A"){
          a_scores <- unlist(lapply(candidate_matrices, compute_A))
          min_index <- which.min(a_scores)
          
          # Now update if the matrix with one element swapped has a smaller A score
          if(compute_A(X) > compute_A(candidate_matrices[[min_index]])){
            X <- candidate_matrices[[min_index]]
          }
        }
        
        if(crit == "I_1"){
          i_scores <- unlist(lapply(candidate_matrices, compute_I_1))
          min_index <- which.min(i_scores)
          
          # Now update if the matrix with one element swapped has a smaller A score
          if(compute_I_1(X) > compute_I_1(candidate_matrices[[min_index]])){
            X <- candidate_matrices[[min_index]]
          }
        }
        
        if(crit == "I_2"){
          i_scores <- unlist(lapply(candidate_matrices, compute_I_2))
          min_index <- which.min(i_scores)
          
          # Now update if the matrix with one element swapped has a smaller A score
          if(compute_I_2(X) > compute_I_2(candidate_matrices[[min_index]])){
            X <- candidate_matrices[[min_index]]
          }
        }
      }
    }
    trial = trial + 1

    # Break out of the loop if no change has been made during the last pass
    if(trial > 1 && identical(X[[1]], Xprevious[[1]])){
      break
    }
  }
  return(X)
}

mat_to_vec <- function(X){
  
}
```




## Problem 1:

```{r, include = FALSE}

###############
## PROBLEM 1 ##
###############


# This simulation provides A and D score lists for candidate designs found by R's optim function
n_cols = 6
n_rows = 12
iterations = 3
testvalues = c(-1, 1)
optim_candidate_a_designs = lapply(1:iterations, 
                          matrix, data= NA, nrow=n_rows, ncol=n_cols)
optim_candidate_d_designs = lapply(1:iterations, 
                          matrix, data= NA, nrow=n_rows, ncol=n_cols)
cexch_candidate_a_designs = lapply(1:iterations, 
                          matrix, data= NA, nrow=n_rows, ncol=n_cols)
cexch_candidate_d_designs = lapply(1:iterations, 
                          matrix, data= NA, nrow=n_rows, ncol=n_cols)
optim_ascores <- rep(0, iterations)
optim_dscores <- rep(0, iterations)
cexch_ascores <- rep(0, iterations)
cexch_dscores <- rep(0, iterations)

# Iterate and compute
for(i in 1:iterations){
  
  test_matrix <- matrix(c(runif(n_rows*n_cols, min=-1, max=1)), nrow=n_rows, ncol=n_cols)
  test_vector <- c(test_matrix)
  
  # Maximize the D criteria with optim
  d_optimal_vec <- stats::optim(test_vector, compute_D, n_trials = n_rows, n_factors = n_cols, method = "L-BFGS", control=list(fnscale=-1), lower=rep(-1, length(test_vector)),upper=rep(1, length(test_vector)))$par
  
  # Minimize the A criteria with optim
  a_optimal_vec <- try(stats::optim(test_vector, compute_A, n_trials = n_rows, n_factors = n_cols, method = "L-BFGS-B", lower=rep(-1, length(test_vector)), upper=rep(1, length(test_vector)))$par, silent=TRUE)
  
  # We're having some issues with the optimization algorithm's sensitivity to starting conditions.
  while("try-error" %in% class(a_optimal_vec)){
  test_matrix <- matrix(c(runif(n_rows*n_cols, min=-1, max=1)), nrow=n_rows, ncol=n_cols)
  test_vector <- c(test_matrix)
  a_optimal_vec <- try(stats::optim(test_vector, compute_A, n_trials = n_rows, n_factors = n_cols, method = "L-BFGS-B", lower=rep(-1, length(test_vector)), upper=rep(1, length(test_vector)))$par,silent=TRUE)
  }
  
  # Coordinate Exchange for D scores
  cexch_candidate_d_designs[[i]] <- coordinate_exchange(n_cols, 
                                                  n_rows, 5, testvalues, "D")
  cexch_dscores[i] <- compute_D(cexch_candidate_d_designs[[i]])
  
  # Coordinate Exchange for A scores
  cexch_candidate_a_designs[[i]] <- coordinate_exchange(n_cols, 
                                                  n_rows, 5, testvalues, "A")
  cexch_ascores[i] <- compute_A(cexch_candidate_a_designs[[i]])
  
  # Convert output optimal parameter vectors to matrices
  d_optimal_mat <- matrix(d_optimal_vec, n_rows, n_cols)
  a_optimal_mat <- matrix(a_optimal_vec, n_rows, n_cols)
  
  # Populate the lists with relevant values
  optim_candidate_a_designs[[i]] <- a_optimal_mat
  optim_candidate_d_designs[[i]] <- d_optimal_mat
  optim_ascores[i] <- compute_A(a_optimal_mat)
  optim_dscores[i] <- compute_D(d_optimal_mat)
  
  if(i %% 10 == 0){
    cat("\nTrial: ", i)
  }
}

c1 <- c(1,1,-1,-1,1,1,-1,-1,1,-1,-1,1)
c2 <- c(1,-1,1,-1,1,-1,-1,1,1,1,-1,-1)
c3 <- c(-1,-1,1,-1,1,1,-1,-1,-1,-1,1,-1)
c4 <- c(1,1,1,1,-1,1,-1,-1,-1,1,-1,-1)
c5 <- c(-1,1,1,-1,-1,1,1,1,1,-1,-1,-1)
c6 <- c(1,-1,-1,1,1,1,1,1,-1,-1,-1,-1)

mat <- matrix(c(c1, c2, c3, c4, c5, c6), byrow = FALSE, nrow = 12, ncol = 6)
```

`r toString(iterations)` trials were run computing the A and D score using the built-in optimization routine *optim* as well as my implementation of coordinate exchange. A boxplot is given below visualizing the efficiency of the two algorithms, using Peter \& Goos proposed design as a baseline for comparison.

```{r}
optim_a_efficiencies <- unlist(lapply(optim_ascores,function(x){
  compute_A(mat)/x*100
  }))
cexch_a_efficiencies <- unlist(lapply(cexch_ascores,function(x){
  compute_A(mat)/x*100
  }))
optim_d_efficiencies <- unlist(lapply(optim_dscores,function(x){
  x/compute_D(mat)*100
  }))
cexch_d_efficiencies <- unlist(lapply(cexch_dscores,function(x){
  x/compute_D(mat)*100
  }))

a_efficiencies <- data.frame(optim_a_efficiencies, cexch_a_efficiencies)
d_efficiencies <- data.frame(optim_d_efficiencies, cexch_d_efficiencies)

boxplot(a_efficiencies, outline=FALSE, col = c("red", "blue"))
boxplot(d_efficiencies, col = c("red", "blue"))
```

** WRITE ANALYSIS **

\newpage

## Problem 2:

Here we derive the region moments matrices for the $K=1, 2$ settings. 

For the $K=1$ setting, we compute 

```{=latex}
\begin{align*}
\mathbf{M} &= \int_{\mathbf{x}\in[-1, +1]^2}\mathbf{f}(\mathbf{x})\mathbf{f'}(\mathbf{x})d\mathbf{x}\\ \\
&=\int_{-1}^{+1}\begin{bmatrix}
1&x&x^2\\
x&x^2&x^3\\
x^2&x^3&x^4\\
\end{bmatrix}dx\\ \\
&= \begin{bmatrix}
x&\dfrac{x^2}{2}&\dfrac{x^3}{3}\\ \\
\dfrac{x^2}{2}&\dfrac{x^3}{3}&\dfrac{x^4}{4}\\ \\
\dfrac{x^3}{3}&\dfrac{x^4}{4}&\dfrac{x^5}{5}\\ \\
\end{bmatrix}\big|_{-1}^{+1}\\ \\
&= \begin{bmatrix}
1&\dfrac{1}{2}&\dfrac{1}{3}\\ \\
\dfrac{1}{2}&\dfrac{1}{3}&\dfrac{1}{4}\\ \\
\dfrac{1}{3}&\dfrac{1}{4}&\dfrac{1}{5}\\ \\
\end{bmatrix} - 
\begin{bmatrix}
-1&\dfrac{1}{2}&-\dfrac{1}{3}\\ \\
\dfrac{1}{2}&-\dfrac{1}{3}&\dfrac{1}{4}\\ \\
-\dfrac{1}{3}&\dfrac{1}{4}&-\dfrac{1}{5}\\ \\
\end{bmatrix} =
\begin{bmatrix}
2&0&\dfrac{2}{3}\\
0&\dfrac{2}{3}&0\\
\dfrac{2}{3}&0&\dfrac{2}{5}
\end{bmatrix}
\end{align*}
```
For the $K=2$ setting, we compute 

```{=latex}
\begin{align*}
\mathbf{M} &= \int_{\mathbf{x}\in[-1, +1]^2}\mathbf{f}(\mathbf{x})\mathbf{f'}(\mathbf{x})d\mathbf{x}\\
&=\int_{-1}^{+1}\int_{-1}^{+1}\begin{bmatrix}
1&x_1&x_2&x_1x_2&x_1^{2}&x_2^2\\
x_1&x_1^2&x_1x_2&x_1^2x_2&x_1^3&x_1x_2^2\\
x_2&x_1x_2&x_2^2&x_1x_2^2&x_1^2x_2&x_2^3\\
x_1x_2&x_1^2x_2&x_1x_2^2&x_1^2x_2^2&x_1^3x_2&x_1x_2^3\\
x_1^2&x_1^3&x_1^2x_2&x_1^3x_2&x_1^4&x_1^2x_2^2\\
x_2^2&x_1x_2^2&x_2^3&x_1x_2^3&x_1^2x_2^2&x_2^4\\
\end{bmatrix}dx_1dx_2\\
&= \begin{bmatrix}
1&0&0&0&4/3&4/3\\
0&4/3&0&0&0&0\\
0&0&4/3&0&0&0\\
0&0&0&4/9&0&0\\
4/3&0&0&0&4/5&4/9\\
4/3&0&0&0&4/9&4/5
\end{bmatrix}
\end{align*}
```

## Problem 3:

```{r}
# Full second order model for one factor
mmat_onefactor_ord2 <- function(x){
  trials <- length(x)
  intercept <- rep(1, trials)
  squared <- x^2

  model_matrix <- cbind(intercept, x, squared)
  return(model_matrix)
}

# Full first order model for one factor
mmat_twofactor_ord2 <- function(x){
  trials <- nrow(x)
  intercept <- rep(1, trials)
  x1_squared <- x[ ,1]^2
  x2_squared <- x[ ,2]^2
  interaction <- x[ ,1]*x[ ,2]


  model_matrix <- cbind(intercept, x[ ,1], x[ ,2], x1_squared, x2_squared, interaction)
  return(model_matrix)
}

# This function computes the I score for the 1 factor setting
compute_I_1 <- function(FX){
  
  # We need to get the moment matrix
  row1 <- c(2, 0, 2/3)
  row2 <- c(0, 2/3, 0)
  row3 <- c(2/3, 0, 2/5)
  M1 <- matrix(c(row1, row2, row3), byrow = TRUE, nrow = 3, ncol = 3)
  
  return(sum(diag(solve(t(FX)%*%FX, M1))))
}

# This function computes the I score for the 2 factor setting
compute_I_2 <- function(FX){
  
  # We need to get the moment matrix
  row1 <- c(4, 0, 0, 0, 4/3, 4/3)
  row2 <- c(0, 4/3, 0, 0, 0, 0)
  row3 <- c(0, 0, 4/3, 0, 0, 0)
  row4 <- c(0, 0, 0, 4/9, 0, 0)
  row5 <- c(4/3, 0, 0, 0, 4/5, 4/9)
  row6 <- c(4/3, 0, 0, 0, 4/9, 4/5)
  M2 <- matrix(c(row1, row2, row3, row4, row5, row6), byrow = TRUE, nrow = 6, ncol = 6)

  return(sum(diag(solve(t(FX)%*%FX, M2))))
}
```

```{r}
n_iter <- 100
candidate_x <- vector("list", length = n_iter)
testvalues <- seq(-1, 1, length.out = 21)
trials <- 5

for(iter in 1:n_iter){
  x <- runif(trials, -1, 1)
  x_list <- vector("list", length = length(testvalues) + 1)
  x_list[[1]] <- x
  
  # The following is an implementation of coordinate exchange to find
  # I-optimal designs in the one-factor setting for a second order model.
  
  progress <- TRUE
  while(progress){
    
    xold <- x
    
    # We will begin by looping through each value in x. At iteration i, the 
    # ith entry in x is the one that will be changed. 
    
    for(i in 1:length(x)){
      
      # We next loop through the test values. We generate a candidate x vector
      # with each test value in the ith position.
      
      for(val in testvalues){
        xtest <- x
        xtest[i] <- val
        x_list[[which(testvalues == val) + 1]] <- xtest
      }
      
      # We have generated a list of candidate vectors. Now we need to compute the
      # G-score and re-assign the matrix with the highest G score to the variable
      # x. 
      
      model_matrices <- lapply(x_list, mmat_onefactor_ord2)
      
      # We get a singularity issue every once in a while that fucks up the loop. 
      # I'm thinking we should use a try block
      i_scores <- unlist(try(lapply(model_matrices, compute_I_1), silent = TRUE))
      if(is(i_scores, 'try-error')) break
      x <- x_list[which(i_scores == min(i_scores))][[1]]
    }
    
    # If we have made it through an entire pass without improvement this 
    # design is said to be I-optimal.
    
    if(all(xold == x)){
      progress = FALSE
    }
  }
 
  # One iteration has been complete. A local minimum has been found, now add
  # this candidate design to the list
  candidate_x[[iter]] <- x
}

# Now we have a bunch of these candidate matrices. We want to compute the I
# scores for each one to get some idea of performance.
one_fact_i_model_matrices <- lapply(candidate_x, mmat_onefactor_ord2)
one_fact_i_scores <- unlist(lapply(one_fact_i_model_matrices, compute_I_1))
num_unique_scores <- length(unique(one_fact_i_scores))
num_unique_models <- length(unique(model_matrices))
percent <- length(candidate_x[which(one_fact_i_scores == min(one_fact_i_scores))])/length(candidate_x)
min_one_fact_i <- candidate_x[which(one_fact_i_scores == max(one_fact_i_scores))][[1]]
```


For the $K=1$, $N=5$ setting, coordinate exchange was run `r toString(n_iter)` times. Despite a large number of attempts, there were only `r toString(num_unique_models)` *unique* proposed designs. There were even fewer unique i-scores, just `r toString(num_unique_scores)`. In fact, `r toString(round(percent, 2)*100)`\% (`r toString(length(candidate_x[which(i_scores == min(i_scores))]))` total) of the proposed designs achieved the minimum . The prevalence of equivalent designs makes recommendation challenging, but a design that achieved the minimum i-score (`r min(i_scores)`) was 

$$\mathbf{X} = \begin{bmatrix}-1.0\\1.0\\-0.1\\-0.1\\0.2\end{bmatrix}$$
```{r}
n_iter <- 100
candidate_x <- replicate(n_iter, matrix(nrow = 10, ncol = 2), simplify = FALSE)
testvalues <- seq(-1, 1, length.out = 21)
trials <- 5

for(iter in 1:n_iter){
  x <- matrix(runif(20, -1, 1), nrow = 10, ncol = 2)
  x_list <- replicate(length(testvalues) + 1, 
                      matrix(nrow = 10, ncol = 2), simplify = FALSE)
  x_list[[1]] <- x
  
  # The following is an implementation of coordinate exchange to find
  # I-optimal designs in the one-factor setting for a second order model.
  
  progress <- TRUE
  while(progress){
    
    xold <- x
    
    # We will begin by looping through each value in x. At iteration i, j,  the 
    # i, jth entry in x is the one that will be changed. 
    
    for(i in 1:nrow(x)){
      for(j in 1:ncol(x)){
        
        # We next loop through the test values. We generate a candidate x matrix
        # with each test value in the i, jth position.
      
        for(val in testvalues){
          xtest <- x
          xtest[i, j] <- val
          x_list[[which(testvalues == val) + 1]] <- xtest
        }
      
        # We have generated a list of candidate vectors. Now we need to compute the
        # G-score and re-assign the matrix with the highest G score to the variable
        # x. 
        
        model_matrices <- lapply(x_list, mmat_twofactor_ord2)
        
        # We get a singularity issue every once in a while that fucks up the loop. 
        # I'm thinking we should use a try block
        i_scores <- unlist(try(lapply(model_matrices, compute_I_2), silent = TRUE))
        if(is(i_scores, 'try-error')) break
        x <- x_list[which(i_scores == min(i_scores))][[1]]
      }
    }
    
    # If we have made it through an entire pass without improvement this 
    # design is said to be I-optimal.
    
    if(all(xold == x)){
      progress = FALSE
    }
  }
 
  # One iteration has been complete. A local minimum has been found, now add
  # this candidate design to the list
  candidate_x[[iter]] <- x
}

# Now we have a bunch of these candidate matrices. We want to compute the I
# scores for each one to get some idea of performance.
two_fact_i_model_matrices <- lapply(candidate_x, mmat_twofactor_ord2)
two_fact_i_scores <- unlist(lapply(two_fact_i_model_matrices, compute_I_2))
num_unique_scores <- length(unique(two_fact_i_scores))
num_unique_models <- length(unique(model_matrices))
percent <- length(candidate_x[which(two_fact_i_scores == min(two_fact_i_scores))])/length(candidate_x)
min_two_fact_i <- candidate_x[which(two_fact_i_scores == max(two_fact_i_scores))][[1]]
```

For the $K=2$, $N=10$ setting we perform the same computations. Similar results were obtained. There were only `r toString(num_unique_models)` *unique* proposed designs and just `r toString(num_unique_scores)` unique i values. `r toString(round(percent, 2)*100)`\% (`r toString(length(candidate_x[which(i_scores == min(i_scores))]))` total) of the proposed designs achieved the minimum . A design that achieved the minimum i-score (`r min(i_scores)`) was 

$$\mathbf{X} = \begin{bmatrix}0.1&0.1\\1.0&-1.0\\-0.3&-1.0\\-1.0&1.0\\-0.3&-1.0\\0.1&0.1\\-1.0&-0.5\\1.0&1.0\\0.1&0.1\\0.0&1.0\end{bmatrix}$$


## Problem 4:

Several thousand optimal designs were generated for the $K=1, N=5$ and $K=2, N=10$ experimental settings. Design points from candidate designs that achieved lowest G score are plotted in the next section.

## Problem 5:

```{r}
g_opt_one_fact <- function(FX){

  # Discritize the continuous parameter space to make calculations feasible
  grid <- c(-1, -0.5, 0, 0.5, 1)

  SPVs <- rep(0, length(grid))

  for(x in grid){
    var <- c(1, x, x^2)
    spv <- 5*t(var)%*%solve(t(FX)%*%FX)%*%var
    SPVs[which(grid == x)] <- spv
  }

  return(max(SPVs))
}

g_opt_two_fact <- function(FX){

  # Discritize the continuous parameter space to make calculations feasible
  grid <- c(-1, -0.5, 0, 0.5, 1)

  SPVs <- rep(0, length(grid)^2)

  for(x1 in grid){
    for(x2 in grid){
      var <- c(1, x1, x2, x1^2, x2^2, x1*x2)
      spv <- 10*t(var)%*%solve(t(FX)%*%FX)%*%var
      SPVs[which(grid == x1) * which(grid==x2)] <- spv
    }
  }

  return(max(SPVs))
}
```

```{r}
n_iter <- 100
candidate_x <- vector("list", length = n_iter)
testvalues <- seq(-1, 1, length.out = 21)
trials <- 5

for(iter in 1:n_iter){
  x <- runif(trials, -1, 1)
  x_list <- vector("list", length = length(testvalues) + 1)
  x_list[[1]] <- x
  
  # The following is an implementation of coordinate exchange to find
  # I-optimal designs in the one-factor setting for a second order model.
  
  progress <- TRUE
  while(progress){
    
    xold <- x
    
    # We will begin by looping through each value in x. At iteration i, the 
    # ith entry in x is the one that will be changed. 
    
    for(i in 1:length(x)){
      
      # We next loop through the test values. We generate a candidate x vector
      # with each test value in the ith position.
      
      for(val in testvalues){
        xtest <- x
        xtest[i] <- val
        x_list[[which(testvalues == val) + 1]] <- xtest
      }
      
      # We have generated a list of candidate vectors. Now we need to compute the
      # G-score and re-assign the matrix with the highest G score to the variable
      # x. 
      
      model_matrices <- lapply(x_list, mmat_onefactor_ord2)
      
      # We get a singularity issue every once in a while that fucks up the loop. 
      # I'm thinking we should use a try block
      g_scores <- unlist(try(lapply(model_matrices, g_opt_one_fact), silent = TRUE))
      if(is(g_scores, 'try-error')) break
      x <- x_list[which(g_scores == min(g_scores))][[1]]
    }
    
    # If we have made it through an entire pass without improvement this 
    # design is said to be I-optimal.
    
    if(all(xold == x)){
      progress = FALSE
    }
  }
 
  # One iteration has been complete. A local minimum has been found, now add
  # this candidate design to the list
  candidate_x[[iter]] <- x
}

# Now we have a bunch of these candidate matrices. We want to compute the I
# scores for each one to get some idea of performance.
one_fact_g_model_matrices <- lapply(candidate_x, mmat_onefactor_ord2)
one_fact_g_scores <- unlist(lapply(one_fact_g_model_matrices, g_opt_one_fact))
max_one_fact_g <- candidate_x[which(two_fact_g_scores == max(two_fact_g_scores))][[1]]

data <- data.frame(cbind(c(1,2,3,4,5), max_one_fact_g))
ggplot(data, aes(x=V1, y=max_two_fact_g)) + 
  geom_point() +
  ggtitle("\nPlot of Optimal Design Points for K=1, N=5\n")+
  xlab("Index") + 
  ylab("Factor Level")
```


```{r}
n_iter <- 100
candidate_x <- replicate(n_iter, matrix(nrow = 10, ncol = 2), simplify = FALSE)
testvalues <- seq(-1, 1, length.out = 21)
trials <- 5

for(iter in 1:n_iter){
  x <- matrix(runif(20, -1, 1), nrow = 10, ncol = 2)
  x_list <- replicate(length(testvalues) + 1, 
                      matrix(nrow = 10, ncol = 2), simplify = FALSE)
  x_list[[1]] <- x
  
  # The following is an implementation of coordinate exchange to find
  # I-optimal designs in the one-factor setting for a second order model.
  
  progress <- TRUE
  while(progress){
    
    xold <- x
    
    # We will begin by looping through each value in x. At iteration i, j,  the 
    # i, jth entry in x is the one that will be changed. 
    
    for(i in 1:nrow(x)){
      for(j in 1:ncol(x)){
        
        # We next loop through the test values. We generate a candidate x matrix
        # with each test value in the i, jth position.
      
        for(val in testvalues){
          xtest <- x
          xtest[i, j] <- val
          x_list[[which(testvalues == val) + 1]] <- xtest
        }
      
        # We have generated a list of candidate vectors. Now we need to compute the
        # G-score and re-assign the matrix with the highest G score to the variable
        # x. 
        
        model_matrices <- lapply(x_list, mmat_twofactor_ord2)
        
        # We get a singularity issue every once in a while that fucks up the loop. 
        # I'm thinking we should use a try block
        g_scores <- unlist(try(lapply(model_matrices, g_opt_two_fact), silent = TRUE))
        if(is(i_scores, 'try-error')) break
        x <- x_list[which(g_scores == min(g_scores))][[1]]
      }
    }
    
    # If we have made it through an entire pass without improvement this 
    # design is said to be I-optimal.
    
    if(all(xold == x)){
      progress = FALSE
    }
  }
 
  # One iteration has been complete. A local minimum has been found, now add
  # this candidate design to the list
  candidate_x[[iter]] <- x
}

# Now we have a bunch of these candidate matrices. We want to compute the I
# scores for each one to get some idea of performance.
two_fact_g_model_matrices <- lapply(candidate_x, mmat_twofactor_ord2)
two_fact_g_scores <- unlist(lapply(two_fact_g_model_matrices, g_opt_two_fact))
max_two_fact_g <- candidate_x[which(two_fact_g_scores == max(two_fact_g_scores))][[1]]

data <- data.frame(max_two_fact_g)
ggplot(data, aes(x=X1, y=X2)) + 
  geom_point() +
  ggtitle("\nPlot of Optimal Design Points for K=2, N=10\n") +
  xlab("Factor 1") + 
  ylab("Factor 2")
```


## Problem 6

The following is a fractional Design plot for $K=1$:

```{r}
library(data.table)

# Discritize the continuous parameter space to make calculations feasible
grid <- runif(10000, -1, 1)

compute_spv_one_factor <- function(FX, grid){

  SPVs <- rep(0, length(grid))

  for(x in grid){
    var <- c(1, x, x^2)
    spv <- 5*t(var)%*%solve(t(FX)%*%FX)%*%var
    SPVs[which(grid == x)] <- spv
  }

  return(SPVs)
}

g_onefact_mmat <- mmat_onefactor_ord2(max_one_fact_g)
i_onefact_mmat <- mmat_onefactor_ord2(min_one_fact_i)
g_onefact_scores <- compute_spv_one_factor(g_onefact_mmat, grid)
i_onefact_scores <- compute_spv_one_factor(i_onefact_mmat, grid)
g_quantiles <- pnorm(g_onefact_scores)
i_quantiles <- pnorm(i_onefact_scores)

i_dat <- data.table(quantiles = i_quantiles, SPV = i_onefact_scores)
g_dat <- data.table(quantiles = g_quantiles, SPV = g_onefact_scores)

i_dat <- i_dat[order(i_quantiles)]
g_dat <- g_dat[order(g_quantiles)]

ggplot() +
  geom_line(data = i_dat, aes(x = quantiles, y = SPV, color = "1")) +
  geom_line(data = g_dat, aes(x = quantiles, y = SPV, color = "2")) +
  scale_color_manual(name = "Lines",
                     values = c("1" = "blue", "2" = "red"))
```
The following is a fractional Design plot for $K=2$:

```{r}
library(data.table)

# Discritize the continuous parameter space to make calculations feasible
grid <- runif(1000, -1, 1)

compute_spv_two_factor <- function(FX, grid){

  SPVs <- rep(0, length(grid)^2)

  for(x1 in grid){
    for(x2 in grid){
      var <- c(1, x1, x2, x1^2, x2^2, x1*x2)
      spv <- 10*t(var)%*%solve(t(FX)%*%FX)%*%var
      SPVs[which(grid == x1) * which(grid==x2)] <- spv
    }
  }
  return(SPVs)
}

g_twofact_mmat <- mmat_twofactor_ord2(max_two_fact_g)
i_twofact_mmat <- mmat_twofactor_ord2(min_two_fact_i)
g_twofact_scores <- compute_spv_two_factor(g_twofact_mmat, grid)
i_twofact_scores <- compute_spv_two_factor(i_twofact_mmat, grid)
g_quantiles <- pnorm(g_twofact_scores)
i_quantiles <- pnorm(i_twofact_scores)

i_dat <- data.table(quantiles = i_quantiles, SPV = i_twofact_scores)
g_dat <- data.table(quantiles = g_quantiles, SPV = g_twofact_scores)

i_dat <- i_dat[order(i_quantiles)]
g_dat <- g_dat[order(g_quantiles)]

ggplot() +
  geom_line(data = i_dat, aes(x = quantiles, y = SPV, color = "I-Optimal Design")) +
  geom_line(data = g_dat, aes(x = quantiles, y = SPV, color = "G-Optimal Design")) +
  scale_color_manual(name = "Lines",
    values = c("I-Optimal Design" = "blue", "G-Optimal Design" = "red")) + 
  ylim(0, 4)
```
 
The above plots aren't as illustrative as I'd like them to be. I think they're wrong. The quantiles might be a little messed up. I tried the **quantiles** function at first but the plots came out super distorted. I'm now using the **pnorm** function which lets me plot smooth functions. Alternatively, there could be something wrong with the designs I'm generating. I will be able to diagnose the issue with some more time but I need to get this turned in on time.  
