---
title: "Problem Space Exploration"
author: "Hyrum Hansen"
date: '2023-04-09'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Math Formula Derivations

Given a one-factor second order model, we derive a formula to compute SPV. Recall the second order model expressed in scalar form:

$$\hat{y} = \beta_0 + \beta_1x_1 + \beta_2x^2 + \epsilon$$
which can be generalized to 

$$\mathbf{\hat{y}} = \mathbf{F}(\mathbf{X})\mathbf{\beta} + \epsilon$$
in matrix-vector form. Now we derive the SPV formula for an arbitrary 3-trial experimental setting. That is, 

$$\mathbf{x} = \begin{bmatrix}a\\b\\c\end{bmatrix}$$
where $a, b$, and $c$ represent factor levels for the univariate experiment. Now the model matrix is 

$$\mathbf{F}(\mathbf{X}) = \begin{bmatrix}1&a&a^2\\1&b&b^2\\1&c&c^2\end{bmatrix}.$$
To compute the SPV for this arbitrary model matrix, we need to compute

$$SPV(\mathbf{x}^T|\mathbf{X}) = Nf^T(\mathbf{x}^T)(\mathbf{F}(\mathbf{X})^T\mathbf{F}(\mathbf{X}))^{-1}f(\mathbf{x})$$
so to make the calculation practicable lets make the inverse information matrix arbitrary:

$$(\mathbf{F}(\mathbf{X})^T\mathbf{F}(\mathbf{X}))^{-1} =\begin{bmatrix}A&B&C\\D&E&F\\G&H&I\end{bmatrix}.$$
Now for the one-factor setting we have 

\begin{align*}
SPV(\mathbf{x}^T|\mathbf{X}) &= 3\times  \begin{bmatrix}1&x&x^2\end{bmatrix}\begin{bmatrix}A&B&C\\D&E&F\\G&H&I\end{bmatrix}\begin{bmatrix}1\\x\\x^2\end{bmatrix}\\ \\
&= 3(A + x(D+B) + x^2(C+E+G) + x^3(H+F) + Ix^4)
\end{align*}

The two factor setting is 

\begin{align*}
SPV(\mathbf{x}^T|\mathbf{X}) &= 3\times  \begin{bmatrix}1&x_1&x_2&x_1^2&x_2^2&x_1x_2\end{bmatrix}\begin{bmatrix}f_{11}&\hdots&f_{16}\\
\vdots&\ddots&\vdots\\f_{61}&\hdots&f_{66}\end{bmatrix}\begin{bmatrix}1\\x_1\\x_2\\x_1^2\\x_2^2\\x_1x_2\end{bmatrix}\\ \\
&= \begin{bmatrix}\vec{x}^T\vec{f}_{1i}&\vec{x}^T\vec{f}_{2i}&\vec{x}^T\vec{f}_{3i}&\vec{x}^T\vec{f}_{4i}&\vec{x}^T\vec{f}_{5i}&\vec{x}^T\vec{f}_{6i}\end{bmatrix}\begin{bmatrix}1\\x_1\\x_2\\x_1^2\\x_2^2\\x_1x_2\end{bmatrix}\\ \\
&= \vec{x}^T\vec{f}_{1i}+\vec{x}^T\vec{f}_{2i}x_1+\vec{x}^T\vec{f}_{3i}x_2+\vec{x}^T\vec{f}_{4i}x_1^2+\vec{x}^T\vec{f}_{5i}x_2^2+\vec{x}^T\vec{f}_{6i}x_1x_2
\end{align*}

The highest order terms in this expression will be $x_1^4$ and $x_2^4$. 

## Plotting SPV

For the K=1 case (and an arbitrary inverse information matrix):

```{r echo = FALSE}
curve(3*(1*x^2 + 0.5*x + -0.5)*x^2 + 3*-0.5*x^2 + 3*(1*x^2 + 0.5*x + 0)*x + 3*1*x + 3*-1, from=-2, to=1, xlab="x", ylab="SPV")
```
Here's a similar plot for the K=2 case. For this problem the candidate design 
matrix is arbitrary.

```{r, echo = FALSE}
# Full second order model for one factor
mmat_twofactor_ord2 <- function(x){
  trials <- nrow(x)
  intercept <- rep(1, trials)
  x1_squared <- x[ ,1]^2
  x2_squared <- x[ ,2]^2
  interaction <- x[ ,1]*x[ ,2]


  model_matrix <- cbind(intercept, x, x1_squared, x2_squared, interaction)
  return(model_matrix)
}
```


```{r, echo = FALSE}
suppressWarnings(library(ggplot2))
suppressWarnings(library(MASS))
suppressWarnings(library(reshape2))
suppressWarnings(library(reshape))

x1 <- sort(runif(100, -1, 1))
x2 <- sort(runif(100, -1, 1))

candidate_x <- matrix(runif(12, -1, 1), 
                      byrow = TRUE, nrow = 6, ncol = 2)
FX <- mmat_twofactor_ord2(candidate_x)
spv <- rep(0, length(x1)*length(x2))

i=0
for(my_x1 in x1){
  for(my_x2 in x2){
    i = i + 1
    var <- c(1, my_x1, my_x2, my_x1^2, my_x2^2, my_x1*my_x2)
    spv[i] <- t(var)%*%solve(t(FX)%*%FX)%*%var
  }
}

spv <- matrix(spv, nrow = length(x1), ncol = length(x2))

mtrx.melt <- suppressWarnings(melt(spv, id.vars = c("x1", "x2"), measure.vars = "SPV"))
names(mtrx.melt) <- c("x1", "x2", "SPV")

suppressWarnings(ggplot(mtrx.melt, aes(x = x1, y = x2, z = SPV)) +
         stat_contour(geom = "polygon", aes(fill = ..level..)) +
         geom_tile(aes(fill = SPV)) +
         stat_contour(bins = 15) +
         xlab("x1") +
         ylab("x2") +
         guides(fill = guide_colorbar(title = "SPV for random design matrix")))
```




